---
title: "ES193 Homework 3 Code"
author: "Jared Umphress"
subtitle: "May 28, 2026"
format: pdf
---

# Part 1. Set up tasks

Load in libraries 
```{r}
#| message: false
#| warning: false

# loading in libraries for this homework
library(tidyverse)
library(here)
library(janitor)
library(readxl)
library(ggplot2)
library(ggeffects)
library(flextable)

```


```{r}
# store temp-kelp.csv as an object called kelp
kelp <- read_csv(here("data", "temp-kelp.csv"))

# store my personal data 
my_data <- read_csv(here("data/screen_time.csv"))

```


# Part 2. Problems
## Problem 1. Giant kelp fronds
a. An appropriate test\newline
The two appropriate tests to determine the strength of the relationship between temperature and giant kelp frond elongation rate are Pearson's correlation and Spearman rank correlation; both of which assume that observations are independent. A Pearson's correlation test is performed for a parametric sample where it is assumed that the variables are continuous, normally distributed, and that there is a linear relationship between them. A Spearman rank correlation test is performed for a non-parametric sample where it is assumed that there is a monotonic relationship between the variables. 


b. Create a visualization\newline

```{r}

# ggplot as base layer 
ggplot(
  # data is kelp
  # temp is x-axis
  # kelp_elongation is y-axis
  data = kelp,
  mapping = aes(x = temp_c,
                y = kelp_elong
                )
) + 
  # geom_point as first layer
  geom_point(
    # changing color of points
    color = 'springgreen4' 
  ) +
  # relabelling x and y axis
  labs(
    x = "Temperatures (Celcius)",
    y = "Kelp Elongation Rate (cm/day)"
  ) +
  # changing the theme from the default
  theme_minimal()

```

c. Check your assumptions and run your test\newline
### Checking assumptions
```{r}

# ggplot as base layer
ggplot(
  # using kelp data
  data = kelp,
  # plotting temperature 
  mapping = aes(sample = temp_c)
) +
  # qq line is first layer
  geom_qq_line() +
  # qq points are second layer
  # changing color of the points
  geom_qq(color = 'tomato4') +
  # setting theme as minimal
  theme_minimal()



# ggplot as base layer 
ggplot(
  # using kelp data
  data = kelp,
  # plotting kelp elongation rate
  mapping = aes(sample = kelp_elong)
) +
  # qq line is first layer
  geom_qq_line() +
  # qq points are second layer
  # changing color of points
  geom_qq(color = 'skyblue3') +
  # setting theme as minimal
  theme_minimal()


```

I checked the assumption that there is a linear relationship between the two variables by looking at the scatterplot of temperature and kelp elongation rate and determined that there is a linear relationship between them. I checked the assumption that the two variables are normally distributed by looking at a qq plot for each of the variables. Since the points are generally around the qq line and are randomly scattered above and below, I determined that there is a normal distribution for both variables. 


### Running Pearson's Correlation Test
```{r}

# running a correlation test
cor.test(
         # testing correlation between kelp elongation rate and temperature 
         kelp$kelp_elong, kelp$temp_c,
         # method tells you what type of pearson's correlatoin
         method = "pearson")

```

d. Results communication\newline


