---
title: "ES193 Homework 3 Code"
author: "Jared Umphress"
subtitle: "May 28, 2026"
format: pdf
---

# Part 1. Set up tasks

Load in libraries 
```{r}
#| message: false
#| warning: false

# loading in libraries for this homework
library(tidyverse)
library(here)
library(janitor)
library(readxl)
library(ggplot2)
library(ggeffects)
library(flextable)

```


```{r}
# store temp-kelp.csv as an object called kelp
kelp <- read_csv(here("data", "temp-kelp.csv"))

# store my personal data 
my_data <- read_csv(here("data/screen_time.csv"))

```


# Part 2. Problems
## Problem 1. Giant kelp fronds
a. An appropriate test\newline
Linear model and correlation

A linear model is one test that can be used to determine the strength of the relationship with its outputted Coefficient of Determination, which describes how much of the variance in giant kelp frond elongation rate can be explained by temperature. A correlation test informs you the strength between temperature and kelp elongation rate, and whether the relationship is positive or negative.

b. Create a visualization\newline

```{r}

# ggplot as first layer 
ggplot(
  # data is kelp
  # temp is x-axis
  # kelp_elongation is y-axis
  data = kelp,
  mapping = aes(x = temp_c,
                y = kelp_elong
                )
) + 
  geom_point(
    color = 'springgreen4'
  ) +
  labs(
    x = "Temperatures (Celcius)",
    y = "Kelp Elongation Rate (cm/day)"
  ) +
  theme_minimal()






```

